Exploring the Impact of Queer Media Characters on LGBTQ+ Psychological Well-Being: The Role of Queer Media Consumption, Perceived Authenticity, and Intersectionality

### 2025-02-11 Manuscript.qmd
Update results of hypo 1

```{r,include=FALSE}
#| label: setup
# Install required packages
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(psych)){install.packages("psych")}
if(!require(knitr)){install.packages("knitr")}
if(!require(kableExtra)){install.packages("kableExtra")}

@@

```{r,include=FALSE}
#| label: read in cleaned and transformed datasets
# Read in the cleaned and transformed dataset
queer.cleaned <- read_csv("data/queer.cleaned.csv")
queer.hypo <- read_csv("data/queer.hypo.csv")
media.long <- read_csv("data/media.long.csv")

@@

```{r,include=FALSE}
#| label: Source data analysis script
source("Source/Data Analysis.R")

@@

```{r, include=FALSE}
#| label: Prepare for inline character display

@@

### Respondent Characteristics
167 respondents were eligible for the data analysis after data cleaning. Respondents ranged in age from `r min(queer.hypo$age)` to `r max(queer.hypo$age)` years old. The average age of the sample was `r round(mean(queer.hypo$age), 2)` years old (SD = `r round(sd_age, 2)`). This wide age range ensured the study captured perspectives from individuals across different generations. Respondents' residential countries included `r n_distinct (queer.hypo.country$country)` countries and four continents (see Fig 1), with both developed and developing country status, providing a global perspective. Regarding sexuality, the sample featured a diverse range of identities, with `r round(100 * sum(queer.hypo$gender_identity == "Man") / nrow(queer.hypo), 1)`% identifying as men, `r round(100 * sum(queer.hypo$gender_identity == "Woman") / nrow(queer.hypo), 1)`% as women, and 10.2% as non-binary, genderfluid, gender non-conforming, or agender. Among the sample, `r round(100 * sum(queer.hypo$transgender == "Yes") / nrow(queer.hypo), 1)`% was identified as transgender. The sexual orientation distribution of participants was also diverse, featuring `r sexual_orientation_inline` (see Fig 2). The racial distribution of the sample covered `r round(100 * sum(queer.hypo$race == "White") / nrow(queer.hypo), 1)`% white respondents, `r round(100 * sum(queer.hypo$race == "Black or African-American") / nrow(queer.hypo), 1)`% black or African-American respondents, and 15.5% Asian, Indian, or Alaska native. This rich diversity in age (generation), residential country, gender identity, sexual orientation, and racial identity enables meaningful comparisons and analyses across various LGBTQ+ subgroups and their intersecting identities.

@@

>>>>

```{r}
#| label: setup
#| include: false
# Install required packages
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(psych)){install.packages("psych")}
if(!require(knitr)){install.packages("knitr")}
if(!require(kableExtra)){install.packages("kableExtra")}

@@

```{r}
#| label: read in cleaned and transformed datasets
#| include: false
# Read in the cleaned and transformed dataset
queer.cleaned <- read_csv("data/queer.cleaned.csv")
queer.hypo <- read_csv("data/queer.hypo.csv")
media.long <- read_csv("data/media.long.csv")

@@

```{r}
#| label: Source data analysis script
#| include: false
source("Source/Data Analysis.R")

@@

```{r}
#| label: Prepare for inline character display
#| include: false

@@

### Respondent Characteristics
167 respondents were eligible for the data analysis after data cleaning. Respondents ranged in age from `r min(queer.hypo$age)` to `r max(queer.hypo$age)` years old. The average age of the sample was `r round(mean(queer.hypo$age), 2)` years old (*SD* = `r round(sd_age, 2)`). This wide age range ensured the study captured perspectives from individuals across different generations. Respondents' residential countries included `r n_distinct (queer.hypo.country$country)` countries and four continents (Figure 1), with both developed and developing country status, providing a global perspective. Regarding sexuality, the sample featured a diverse range of identities, with `r round(100 * sum(queer.hypo$gender_identity == "Man") / nrow(queer.hypo), 1)`% identifying as men, `r round(100 * sum(queer.hypo$gender_identity == "Woman") / nrow(queer.hypo), 1)`% as women, and 10.2% as non-binary, genderfluid, gender non-conforming, or agender. Among the sample, `r round(100 * sum(queer.hypo$transgender == "Yes") / nrow(queer.hypo), 1)`% was identified as transgender. The sexual orientation distribution of participants was also diverse, featuring `r sexual_orientation_inline` (Figure 2). The racial distribution of the sample covered `r round(100 * sum(queer.hypo$race == "White") / nrow(queer.hypo), 1)`% white respondents, `r round(100 * sum(queer.hypo$race == "Black or African-American") / nrow(queer.hypo), 1)`% black or African-American respondents, and 15.5% Asian, Indian, or Alaska native. This rich diversity in age (generation), residential country, gender identity, sexual orientation, and racial identity enables meaningful comparisons and analyses across various LGBTQ+ subgroups and their intersecting identities.

@@

```{r}
#| label: ANOVA tests for psychological well-being scores by different LGBTQ+ subgroups
#| include: false

# 1. Age
anova_age_effect <- aov(psychological_wellbeing_score ~ generation, data = queer.hypo)
summary(anova_age_effect)

# p = .70, no significant difference has been found among different generations

# 2. Residential Country Development Status
anova_country_effect <- aov(psychological_wellbeing_score ~ development_status, data = queer.hypo)
summary(anova_country_effect)
summary_country_effect <- summary(anova_country_effect)[[1]]

# p = .014, F = 6.16, significant difference has been found between developed and developing countries

# 3. Gender identity
anova_gender_identity_effect <- aov(psychological_wellbeing_score ~ gender_identity_dual, data = queer.hypo)
summary (anova_gender_identity_effect)
summary_gender_effect <- summary(anova_gender_identity_effect)[[1]]

# p = .11, no significant difference has been found among different gender identity, but there was a directional effect

# 4. Transgender status
anova_transgender_effect <- aov(psychological_wellbeing_score ~ transgender, data = queer.hypo)
summary (anova_transgender_effect)

# p = .25, no significant difference has been found among different gender identity

# 5. Sexual orientation
anova_sexual_orientation_effect <- aov(psychological_wellbeing_score ~ sexual_orientation, data = queer.hypo)
summary (anova_sexual_orientation_effect)
summary_orientation_effect <- summary(anova_sexual_orientation_effect)[[1]]

# p = 0.1, no significant differences have been found among different sexual orientation, but there was directional effect

# 6. Race
anova_race_effect <- aov(psychological_wellbeing_score ~ race_category, data = queer.hypo)
summary(anova_race_effect)
summary_race_effect <- summary(anova_race_effect)[[1]]

# p = .09, marginally significant difference has been found between white and other

One-way ANOVA results showed that there were no significant differences or directional effects in queer media characters’ psychological impacts among different generations (*p* = .70) or between transgenders and non-transgenders (*p* = .25). However, significant differences were found between developed and developing countries' residents (Figure 3). Specifically, residents in developed countries showed lower psychological impacts scores by queer media characters than those in developing countries ($M_{\text{developed}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$development_status == "Developed"]), 2)`$ vs. $M_{\text{developing}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$development_status == "Developing"]), 2)`$; *F*(`r summary_country_effect$Df[1]`, `r summary_country_effect$Df[2]`) = `r round(summary_country_effect$"F value"[1], 2)`; *p* = `r sprintf("%.3f", summary_country_effect$"Pr(>F)"[1])`).

Though the ANOVA result did not show significant differences among different gender identities, there was a directional effect indicated by the violin plot (Figure 4). Specifically, men showed lower psychological impacts scores than women and other gender identities ($M_{\text{men}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$gender_identity_dual == "Man"]), 2)`$ vs. $M_{\text{womenother}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$gender_identity_dual == "Women, or nonbinary, or agender"]), 2)`$; *F*(`r summary_gender_effect$Df[1]`, `r summary_gender_effect$Df[2]`) = `r round(summary_gender_effect$"F value"[1], 2)`; *p* = `r sprintf("%.3f", summary_gender_effect$"Pr(>F)"[1])`). The distribution also indicated that men have more low scores than their women or nonbinary counterparts. Regarding sexual orientation, the ANOVA result did not show significant differences among different sexual orientations, but there was also a directional effect (Figure 5). Specifically, bisexual individuals showed the lowest psychological impacts scores among all sexual orientations, while gay or lesbian had the highest impacts scores (*F*(`r summary_orientation_effect$Df[1]`, `r summary_orientation_effect$Df[2]`) = `r round(summary_orientation_effect$"F value"[1], 2)`; *p* = `r sprintf("%.3f", summary_orientation_effect$"Pr(>F)"[1])`).

```{r}
#| label: fig-plot-country-psych-impact
#| fig-cap: Country Comparison of Queer Characters' Psychological Impacts
#| fig-width: 9
#| fig-height: 5

# Define the theme
theme.clean <- function(){
  theme_bw()+
    theme(axis.text.x = element_text(size = 12, vjust = 1, hjust = 1),
          axis.text.y = element_text(size = 12),
          axis.title.x = element_text(size = 14, face = "plain"),             
          axis.title.y = element_text(size = 14, face = "plain"),
          panel.grid.major.x = element_blank(),                                          
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),  
          plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
          plot.title = element_text(size = 16, vjust = 1, hjust = 0.5),
          plot.caption = element_text (size = 8,vjust = 1, hjust = 0.9),
          legend.text = element_text(size = 12, face = "italic"),          
          legend.position = "right")}

# Visualization: Boxplot

country_effect_boxplot <- ggplot(queer.hypo, aes(development_status, psychological_wellbeing_score)) + 
  geom_boxplot(aes(fill = development_status)) +
  theme_bw() +
  scale_fill_manual(values = c("#dbf1e5", "#4fba7f")) +               
  scale_colour_manual(values = c("#dbf1e5", "#4fba7f")) + 
  labs(title = "Country Comparison of Queer Characters' Psychological Impacts", 
       x = "Development Status", y = "Queer Characters' Psychological Impact Score") +
  theme.clean()+
  theme(axis.text.x = element_text(hjust = 0.5))+
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 3, shape = 18, 
               position = position_dodge(0.75), aes(shape = "Mean"), 
               show.legend = FALSE) +  
  stat_summary(fun = mean, geom = "text", aes(label = round(after_stat(y), 2)), 
               color = "black", vjust = -1, position = position_dodge(0.75), 
               size = 4, show.legend = FALSE) +
  labs(fill = "Development Status")

print (country_effect_boxplot)

```{r}
#| label: fig-plot-gender-psych-impact
#| fig-cap: Gender Identity Difference of Queer Characters' Psychological Impacts
#| fig-width: 10
#| fig-height: 5

# Visualization: Violinplot

gender_identity_effect_violinplot <- ggplot(queer.hypo, aes(gender_identity_dual, psychological_wellbeing_score)) + 
  geom_violin(aes(fill = gender_identity_dual)) +
  theme_bw() +
  scale_fill_manual(values = c("#6fa8dc", "#d5a6bd")) +               
  scale_colour_manual(values = c("#6fa8dc", "#d5a6bd")) + 
  labs(title = "Gender Identity Difference of Queer Characters' Psychological Impacts", 
       x = "Gender Identity", y = "Queer Characters' Psychological Impact Score") +
  theme.clean()+
  theme(axis.text.x = element_text(hjust = 0.5))+
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  stat_summary(fun = mean, geom = "point", color = "white", size = 3, shape = 18, 
               position = position_dodge(0.75), aes(shape = "Mean"), 
               show.legend = FALSE) +  
  stat_summary(fun = mean, geom = "text", aes(label = round(after_stat(y), 2)), 
               color = "white", vjust = -1, position = position_dodge(0.75), 
               size = 4, show.legend = FALSE) +
  labs(fill = "Gender Identity")

print (gender_identity_effect_violinplot)

![Sexual Orientation Comparison of Queer Characters' Psychological Impacts](Image/Fig 10_Orientation_Psych_box.png)

Racial differences in queer media characters’ psychological impacts were marginally significant (Figure 6). Specifically, white respondents showed lower psychological impacts scores than other racial groups ($M_{\text{white}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$race_category == "White"]), 2)`$ vs. $M_{\text{blackother}} = `r round(mean(queer.hypo$psychological_wellbeing_score[queer.hypo$race_category == "Black, Asian, Indian, and Others"]), 2)`$; *F*(`r summary_race_effect$Df[1]`, `r summary_race_effect$Df[2]`) = `r round(summary_race_effect$"F value"[1], 2)`; *p* = `r sprintf("%.3f", summary_race_effect$"Pr(>F)"[1])`).

```{r}
#| label: fig-plot-race-psych-impact
#| fig-cap: Race Comparison of Queer Characters' Psychological Impacts
#| fig-width: 10
#| fig-height: 5
# Visualization: Violinplot
race_effect_boxplot <- ggplot(queer.hypo, aes(race_category, psychological_wellbeing_score)) + 
  geom_boxplot(aes(fill = race_category)) +
  theme_bw() +
  scale_fill_manual(values = c("#cfb5a0","#f1e2c8")) +               
  scale_colour_manual(values = c("#cfb5a0","#f1e2c8")) + 
  labs(title = "Race Comparison of Queer Characters' Psychological Impacts", 
       x = "Racial Identity", y = "Queer Characters' Psychological Impact Score") +
  theme.clean()+
  theme(axis.text.x = element_text(hjust = 0.5))+
  scale_y_continuous(breaks = seq(floor(min(queer.hypo$psychological_wellbeing_score)), 
                                  ceiling(max(queer.hypo$psychological_wellbeing_score)), by = 1)) +
  stat_summary(fun = mean, geom = "point", color = "white", size = 3, shape = 18, 
               position = position_dodge(0.75), aes(shape = "Mean"), 
               show.legend = FALSE) +  
  stat_summary(fun = mean, geom = "text", aes(label = round(after_stat(y), 2)), 
               color = "white", vjust = -1, position = position_dodge(0.75), 
               size = 4, show.legend = FALSE) +
  labs(fill = "Racial Identity")

print (race_effect_boxplot)
